---
title: "Conduct rough estimates on waveform LiDAR data with waveformlidar"
author: "Tan Zhou"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Conduct rough estimates on waveform LiDAR data with waveformlidar}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
params:
  EVAL: !r identical(Sys.getenv("NOT_CRAN"), "true")
---

```{r, SETTINGS-knitr, include=FALSE}
stopifnot(require(knitr))
options(width = 90)
opts_chunk$set(
  comment = NA,
  message = FALSE,
  warning = FALSE,
  eval = if (isTRUE(exists("params"))) params$EVAL else FALSE,
  dev = "png",
  dpi = 150,
  fig.asp = 0.8,
  fig.width = 5,
  out.width = "60%",
  fig.align = "center"
)
library(waveformlidar)
#ggplot2::theme_set(theme_default())
```

## Goal

The purpose of this vignette is to prepare the waveform LiDAR data processing using **waveformlidar** package. In addition, we also demonstate an example using **peakfind** function to quickly identify the peak(s) of waveforms.


## A Case Study

```{r prepare formula for decomposition}

##for known estimated paramters using Gaussian model, these three should have the same length
A<-c(76,56,80); u<-c(29,40,67); sig<-c(4,3,2.5) 

fg<-gennls(A,u,sig)
fg$formula

##using adaptive Gaussian model, we still assign the initiated r to 2, you also can give another reasonble value
A<-c(76,56,80); u<-c(29,40,67); sig<-c(4,3,2.5);  r<-c(2,2,2) 
afg<- agennls(A,u,sig,r)
afg$formula
```


###Prepare data and obtain rough estimates of waveforms
```{r rough estimates}
###rough estimate without decompostion or deconvolution
library(data.table)
library(caTools)

data(return)
wf<-data.table(index=c(1:nrow(return)),return)

num_wf<- npeaks(wf[1,], drop=c(1,1))  
num_wf1<- npeaks(return[1,])

rough_estimates<- peakfind(wf[182,])
rough_estimates1<-peakfind(wf[182,], thres = 0.3)

```
The **peakfind** can help us to quickly get the shape of waveforms quickly without any need of implementing complex waveform processing algorithms. The caution should be exercised on results of this function since they are vulnerable to the noise. 

To visually inspect the results, we plot two waveforms as an example.

```{r peakfind to get an overview of the waveform}
y0<- return[182,]
y0[y0==0]<-NA
y1<- as.numeric(y0 - min(y0,na.rm = T) +1)
y1<-y1[1:wavelen(y1)]

realind<-rough_estimates[,3]
newpeak<-rough_estimates[,2]

I<-c("A1","A2","A3","A4");U<-c("u1","u2","u3","u4");S<-c("s1","s2","s3","s4")


#########fig2.b
re<-decom(wf[182,],smooth=TRUE, width=5)[[2]]
yi<-cbind(re[1:3,2],re[4:6,2],re[7:9,2])
sumyi<-0
x<-1:wavelen(y1)
for (i in 1:3){
  sumyi<-sumyi + yi[i,1] * exp(-(x - yi[i,2])**2/(2 * yi[i,3]**2))
}
#re<-decom(wf[2,],smooth=TRUE, width=7)[[3]]
#yi<-cbind(re[1:3,2],re[4:6,2],re[7:9,3])

###fig2.c
y2<- return[2,]
y2[y2==0]<-NA
y3<- as.numeric(y2 - min(y2,na.rm = T) +1)
y3<-y3[1:wavelen(y3)]

rough_estimates2<-peakfind(wf[2,])
realind2<-rough_estimates2[,3]
newpeak2<-rough_estimates2[,2]

I2<-c("A1","A2");U2<-c("u1","u2");S2<-c("s1","s2")

###fig4.d
yi2<-decom(wf[2,])[[3]]
#yi2<-cbind(re2[1:3,2],re[4:6,2],re[7:9,2])
sumyi2<-0
x<-1:wavelen(y2)
for (i in 1:nrow(yi2)){
  sumyi2<-sumyi2 + yi2[i,2] * exp(-(x - yi2[i,3])**2/(2 * yi2[i,4]**2))
}


setwd("A:/research/2018/waveform_summary_R/waveformlidar_package_papaer/code_for_paper") ##change your path here

par(family = 'serif')
png('peakfind_figure2_new.png', width = 7.25*1.8, height = 5.73*1.8, units = 'in',res = 600)
par(mfrow=c(2,2), oma=c(0,0,0,0),mai=c(1.0,1.0,0.8,0))
###fig2.a
plot(seq_along(y1),y1,type="l",col="red",xlab="(a) Time(ns)",ylab="",cex.lab=2.5,lwd=3,axes=F,main="Rough estimates 4 waveform components",cex.main=2)
axis(side=1, at=seq(0,(wavelen(y1)+15),15),cex.axis=2.5)
axis(side=2, at=seq(0,(max(y1)+20),20),cex.axis=2.5)
##add the point peak
xy<-c(0,realind)###for showing sigma
labelxy<-runmean(xy,2)

for (i in 1:length(realind)){
  points(realind[i],newpeak[i],pch=1,col="green",cex=2.5,lwd=2)
  segments(x0=realind[i],y0=0,x1=realind[i],y1=newpeak[i],col="blue",lty=3,lwd=3)##for the vertical
  segments(x0=xy[i],y0=6+(i-1)*7,x1=xy[i+1],y1=6+(i-1)*7,col="black",lty=3,lwd=3)####for the sigma
  text(realind[i],newpeak[i],I[i],adj=1.5,cex=2,col="green") ###for intensity
  text(realind[i],4,U[i],adj=1,cex=2,col="blue") ###for time bin
  text(labelxy[i+1],7*i+2,S[i],adj=0.5,cex=2,col="black")
}
#####fig2.b
##to demonstarte the noise and Gaussian funtion is good for decomposition
x<-1:wavelen(y1)
plot(seq_along(y1),y1,type="l",col="red",xlab="(b) Time(ns)",ylab="Intensity",
     cex.lab=2.5,lwd=3.5,axes=F,main="Unreasonable decomposition",cex.main=2,ylim=c(-30,230))
axis(side=1, at=seq(0,(wavelen(y1)+15),15),cex.axis=2.5)
axis(side=2, at=seq(-30,(max(y1)+40),20),cex.axis=2.5)
lines(seq_along(y1),y1,type="l",col="red",lwd=3)
###idea components
for (i in 1:3){
  lines(x,yi[i,1] * exp(-(x - yi[i,2])**2/(2 * yi[i,3]**2)),lty=2,col=i+2,lwd=3)
}

lines(x,sumyi,lty=3,col="black",lwd=3)
legend("topright",legend=c("1","2","3"),col=c(3,4,5),lty=c(2,2,2),lwd=3,cex=2,box.lwd="white")
legend("topleft",legend=c("IGW","RW"),col=c("black","red"),lty=c(2,1),lwd=3,cex=2,box.lwd="white")
###fig2.c

plot(seq_along(y3),y3,type="l",col="red",xlab="(c) Time(ns)",ylab="",cex.lab=2.5,lwd=3,axes=F,main="Rough estimates 2 waveform components",cex.main=2)
axis(side=1, at=seq(0,(wavelen(y1)+15),15),cex.axis=2.5)
axis(side=2, at=seq(0,(max(y3)+20),20),cex.axis=2.5)
##add the point peak
xy2<-c(0,realind2)###for showing sigma
labelxy2<-runmean(xy2,2)

for (i in 1:length(realind2)){
  points(realind2[i],newpeak2[i],pch=1,col="green",cex=2.5,lwd=2)
  segments(x0=realind2[i],y0=0,x1=realind2[i],y1=newpeak2[i],col="blue",lty=3,lwd=3)##for the vertical
  segments(x0=xy2[i],y0=6+(i-1)*7,x1=xy2[i+1],y1=6+(i-1)*7,col="black",lty=3,lwd=3)####for the sigma
  text(realind2[i],newpeak2[i],I[i],adj=1.5,cex=2,col="green") ###for intensity
  text(realind2[i],4,U[i],adj=1,cex=2,col="blue") ###for time bin
  text(labelxy2[i+1],7*i+2,S2[i],adj=0.5,cex=2,col="black")
}

#####fig2.d
##to demonstarte the noise and Gaussian funtion is good for decomposition
x2<-1:wavelen(y3)
plot(seq_along(y3),y3,type="l",col="red",xlab="(d) Time(ns)",ylab="Intensity",
     cex.lab=2.5,lwd=3.5,axes=F,main="Successful decompostion",cex.main=2,ylim=c(0,max(y3)+20))
axis(side=1, at=seq(0,(wavelen(y3)+15),15),cex.axis=2.5)
axis(side=2, at=seq(-30,(max(y3)+40),20),cex.axis=2.5)
#lines(seq_along(y3),y3,type="l",col="red",lwd=3)
###idea components
for (i in 1:2){
  lines(x2,yi2[i,2] * exp(-(x2 - yi2[i,3])**2/(2 * yi2[i,4]**2)),lty=2,col=i+2,lwd=3)
}

lines(x2,sumyi2,lty=3,col="black",lwd=3)
legend("topright",legend=c("1","2"),col=c(3,4),lty=c(2,2),lwd=3,cex=2,box.lwd="white")
legend("topleft",legend=c("IGW","RW"),col=c("black","red"),lty=c(2,1),lwd=3,cex=2,box.lwd="white")

dev.off()

```
